---
title: "Import and QC of OTG Data"
author: Mike Ackerman
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{otg-import-qc}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
# knitr options
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  comment = "#>"
)

```

```{r setup}
# load necessary libraries
library(DASH)
library(tibble)
library(janitor)
library(dplyr)
library(magrittr)

```

# Introduction

This vignette describes how to import and QC on-the-ground (OTG) habitat data collected using the DASH protocol and Survey123 data collection forms. All of the functionality described here is available as part of the `DASH` R package available [here](https://github.com/BiomarkABS/DASH).

# Data Import

```{r otg-path}
# to your desktop
otg_path = "C:/Users/username/Desktop/1_formatted_csvs/"

# e.g., Biomark NAS
# set NAS prefix, depending on your operating system
if(.Platform$OS.type != 'unix') {
  nas_prefix = "S:"
} else if(.Platform$OS.type == 'unix') {
  nas_prefix = "~/../../Volumes/ABS"
}

# path to otg data
otg_path = paste0(nas_prefix, "/data/habitat/DASH/OTG/2019/lemhi/1_formatted_csvs/")

```

```{r read-otg-csv, eval = F}
# read in just one type of OTG data at a time
otg_raw_cu = read_otg_csv(path = otg_path,
                          otg_type = "CU_1.csv")

otg_raw_wood = read_otg_csv(path = otg_path,
                            otg_type = "Wood_2.csv")

```

```{r read-otg-csv-wrapper, eval = F}
# "loop over" all data types using read_otg_csv_wrapper()
otg_raw = read_otg_csv_wrapper(path = otg_path,
                               otg_type_list = c("surveyPoint_0.csv",
                                                   "CU_1.csv",
                                                   "Wood_2.csv",
                                                   "Jam_3.csv",
                                                   "Undercut_4.csv",
                                                   "Discharge_5.csv",
                                                   "DischargeMeasurements_6.csv"),
                                 otg_type_names = c("survey",
                                                    "cu",
                                                    "wood",
                                                    "jam",
                                                    "undercut",
                                                    "discharge",
                                                    "discharge_measurements"))

```

```{r view-raw-otg}
head(otg_raw$cu)

```

```{r save-raw-otg, eval = F}
save(otg_raw,
     file = paste0(nas_prefix, "/data/habitat/DASH/OTG/2019/lemhi/prepped/otg_raw.rda"))

```

# Initial QC

```{r fake-load-raw-otg, eval = F}
load(paste0(nas_prefix,
            "/data/habitat/DASH/OTG/2019/lemhi/prepped/otg_raw.rda"))

```

```{r init-qc, eval = F}
# QC one data type at a time
init_qc_cu = qc_cu(qc_df = otg_raw$cu)

init_qc_wood = qc_wood(qc_df = otg_raw$wood)

# join qc results together
init_qc_some = init_qc_cu %>%
  add_column(source = "CU",
             .before = 0) %>%
  bind_rows(init_qc_wood %>%
              add_column(source = "Wood",
                         .before = 0))

```

```{r init-qc-all}
# same as above
init_qc_some = qc_wrapper(cu_df = otg_raw$cu,
                          wood_df = otg_raw$wood,
                          redirect_output = FALSE)

# QC all OTG data types at once
init_qc = qc_wrapper(survey_df = otg_raw$survey,
                     cu_df = otg_raw$cu,
                     wood_df = otg_raw$wood,
                     jam_df = otg_raw$jam,
                     undercut_df = otg_raw$undercut,
                     discharge_df = otg_raw$discharge,
                     disch_meas_df = otg_raw$discharge_measurements,
                     redirect_output = FALSE)

```

# Examine QC Results

```{r examine-qc}
tabyl(init_qc,
      source)

init_qc %>%
  print()

```

```{r max-depth-errors}
init_qc %>%
  filter(source == "CU") %>%
  filter(grepl("Column Maximum Depth", error_message)) %>%
  left_join(otg_raw$cu) %>%
  select(source:error_message, `Channel Unit Type`, `Maximum Depth (m)`)
  
```

```{r ocular-ests-errors}
init_qc %>%
  filter(source == "CU") %>%
  filter(grepl("Ocular estimates", error_message)) %>%
  left_join(otg_raw$cu) %>%
  select(source:GlobalID,
         `Channel Unit Type`,
         `Sand/Fines 2mm`:`Boulder 256mm`) %>%
  mutate(sum_ocular = rowSums(.[5:8], na.rm = T))
  
```

```{r fish-cover-errors}
init_qc %>%
  filter(source == "CU") %>%
  filter(grepl("Cover values", error_message)) %>%
  left_join(otg_raw$cu) %>%
  select(source:GlobalID,
         `Overhanging Cover`:`Total No Cover`) %>%
  mutate(sum_cover = rowSums(.[4:8], na.rm = T))
  
```

```{r undercut-errors}
init_qc %>%
  filter(source == "Undercut") %>%
  left_join(otg_raw$undercut) %>%
  select(path_nm:error_message,
         `Undercut Number`:`Width 75% (m)`)
  
```

```{r save-qc, eval = F}
# write QC results to .csv
write_csv(init_qc,
          paste0(nas_prefix, "/data/habitat/DASH/OTG/2019/lemhi/1_formatted_csvs/init_qc.csv"))

```

# Resolving Errors

# "Final" QC

# Data Wrangling and Export

#### END IMPORT AND QC VIGNETTE

<!-- *Holy Guacamole! That deserves a beer!* -->
